Temporal Activity Detection in Untrimmed Videos with Recurrent Neural Networks

### 2016-10-15 paper.tex
Finishing paper

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{amsmath}        % equations
\usepackage{color}

@@

    %In the later years the applications of deep learning techniques have moved from images to videos due to the increase of computational resources available.

    This work proposes a simple pipeline to classify and temporally localize activities in untrimmed videos. Our system uses features from a 3D Convolutional Neural Network (3D-CNN) as input to train a a recurrent neural network (RNN) that learns to classify video clips of 16 frames. After clip prediction, we post-process the output of the RNN to assign a single activity label to each video, and determine the temporal boundaries of the activity within the video. We show how our system can achieve competitive results in both tasks with a simple architecture. We evaluate our method in the ActivityNet Challenge 2016, achieving a 0.00 mAP and a 0.00 mAP in the classification and detection tasks, respectively. Our code and models are publicly acailable at at: \url{ https://github.com/imatge-upc/activitynet-2016-cvprw}

%AMAIA: s'han d'afegir els valors de mAP (o la metrica que sigui)
\end{abstract}

@@

Impressive progress has been reported in the recent literature for video classification \cite{}, which requires to assign a label for the input video. While this task is already challenging, it typically requires videos to be trimmed beforehand (i.e. the category is present in the entire video). However, a video classification system should be able to recognize activities in untrimmed videos, and find the temporal segments in which they appear. This second challenge has been recently proposed in the ActivityNet Challenge 2016 \cite{}, in which participants are asked to both provide a single activity for each video, as well as the temporal segment where the activity happened in the video. % AMAIA: Amb això et refereixes que la tasca de detecció temporal es nova d'Activitynet? Ho he modificat una mica... No queda massa clar... Si és així, jo directament mencionaria la challenge i la citaria.
In order to face both these challenges at the same time, we propose a simple pipeline composed of a 3D-CNN that exploits spatial and short temporal correlations followed by a recurrent neural network which exploits long temporal correlations.

@@

For temporal activity detection, recent works have proposed the usage of Long Short-Term Memory units (LSTM) \cite{}. LSTMs are a type of RNNs that is able to better exploit long and short temporal correlations in sequences, which makes them suitable for video applications. LSTMs have been used alongside CNNs for video classification ~\cite{yao2015describing} and activity localization in videos~\cite{yeung2015every}.

@@

%\paragraph{3D-CNN.}
We use the 3D-CNN model proposed in~\cite{tran2014learning} to extract features for all videos in the database. We split the videos in 16-frame clips and resize them to 171$\times$128 to fit the original model's desired input. Features from the second fully connected layer (fc6) are extracted for each video clip.

@@

%\paragraph{Recurrent Neural Network.}
Next, we design a network that takes a sequence of 3D-CNN-f6 features from a video, and returns a sequence of class probabilities for each element in the sequence. We use LSTM layers, followed with dropout with $p = .5$ and a fully connected layer with a softmax activation. Figure \ref{} shows the proposed architecture. Different configurations of the number of LSTM layers $M$ and the number of cells $c$ have been tested and will be reviewed in the following sections.

@@

\section{Post-Processing}
Given a video, the prediction of our model is sequence of class probabilities for each 16-frame video clip. From this output, we need to be able to predict the activity  and temporally localize it. We post-process the network's output to achieve that goal. First, to obtain the activity prediction for the whole video, we simply compute the mean over all video clips in the video. We then take the class with maximum predicted probability as the predicted class.

@@

\subsection{Dataset}
For all our experiments we use the dataset provided in the ActivityNet Challenge 2016 \cite{}. This dataset contains 640 hours of video and 64 million frames. The ActivityNet dataset is composed of untrimmed videos, providing temporal annotations for the given ground truth class labels.

\subsection{Training}
% AMAIA: He mogut la part de training a experiments
We train the network described in Section \ref{} with the negative log likelihood loss, assigning a lower weight to background samples to deal with dataset imbalance( see equation \ref{eq:loss}).

@@

where $q$ is the predicted probability distribution and $p$ the ground truth probability distribution. In our experiments, we set $\rho$ = X. 
% What is the value of \rho ??

% AMAIA: Aquí tambe hauries d'afegir el numero de iteracions (o epochs), learning rate, optimizer, batch size. Si el tens, el temps d'entrenament en la GPU que entrenessis.

@@

Talk about different values of N and c
talk about different experiments with values of $\gamma$ and k

% AMAIA: Crec que amb una secció on expliquis tots els experiments ja n'hi ha prou. Li he posat el nom de 'Results'

% AMAIA: Si aqui vols parlar del bias towards sport categories, hauries d'afegir la figura que tens que ho demostra. 
\section{Conclusion}

Explain conclusions. Give at the footnote to the repo where all the code is available.

% AMAIA: La url del repo ja l'he posat a l'abstract.

>>>>

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{amsmath}        % equations
\usepackage{color}
\usepackage{graphicx}
\usepackage{subcaption}

@@

    This work proposes a simple pipeline to classify and temporally localize activities in untrimmed videos. Our system uses features from a 3D Convolutional Neural Network (3D-CNN) as input to train a a recurrent neural network (RNN) that learns to classify video clips of 16 frames. After clip prediction, we post-process the output of the RNN to assign a single activity label to each video, and determine the temporal boundaries of the activity within the video. We show how our system can achieve competitive results in both tasks with a simple architecture. We evaluate our method in the ActivityNet Challenge 2016, achieving a 0.5874 mAP and a 0.2237 mAP in the classification and detection tasks, respectively. Our code and models are publicly available at at: \url{https://github.com/imatge-upc/activitynet-2016-cvprw}

%AMAIA: s'han d'afegir els valors de mAP (o la metrica que sigui).
%ALBERTO: FET
\end{abstract}

@@

Impressive progress has been reported in the recent literature for video classification~\cite{tran2014learning,tran2015deep,wang2015towards,yao2015describing}, which requires to assign a label for the input video. While this task is already challenging, it typically requires videos to be trimmed beforehand (i.e. the category is present in the entire video).
However, a video classification system should be able to recognize activities in untrimmed videos, and find the temporal segments in which they appear.
This second challenge has been recently proposed in the ActivityNet Challenge 2016~\cite{caba2015activitynet}, in which participants are asked to both provide a single activity for each video, as well as the temporal segment where the activity happened in the video.
% AMAIA: Amb això et refereixes que la tasca de detecció temporal es nova d'Activitynet? Ho he modificat una mica... No queda massa clar... Si és així, jo directament mencionaria la challenge i la citaria.
In order to face both these challenges at the same time, we propose a simple pipeline composed of a 3D-CNN that exploits spatial and short temporal correlations followed by a recurrent neural network which exploits long temporal correlations.

@@

For temporal activity detection, recent works have proposed the usage of Long Short-Term Memory units (LSTM)~\cite{hochreiter1997long}.
LSTMs are a type of RNNs that is able to better exploit long and short temporal correlations in sequences, which makes them suitable for video applications.
LSTMs have been used alongside CNNs for video classification~\cite{yao2015describing} and activity localization in videos~\cite{yeung2015every}.

@@

We use the 3D-CNN model proposed in~\cite{tran2014learning} to extract features for all videos in the database. We split the videos in 16-frame clips and resize them to 171$\times$128 to fit the original model's desired input. Features from the second fully connected layer (fc6) are extracted for each video clip.

@@

\subsection{Recurrent Neural Network}
\label{sec:architecture}
Next, we design a network that takes a sequence of 3D-CNN-f6 features from a video, and returns a sequence of class probabilities for each element in the sequence. We use LSTM layers, followed with dropout with $p = .5$ and a fully connected layer with a softmax activation. Figure~\ref{fig:global_pipeline} shows the proposed architecture. Different configurations of the number of LSTM layers and the number of cells have been tested and will be reviewed in the following sections.

\begin{figure}[ht]
\centering
\includegraphics[width=.5\linewidth]{img/schematic_pipeline}
\caption{Global architecture of the proposed pipeline.}
\label{fig:global_pipeline}
\end{figure}

@@

\subsection{Post-Processing}
Given a video, the prediction of our model is sequence of class probabilities for each 16-frame video clip. From this output, we need to be able to predict the activity  and temporally localize it. We post-process the network's output to achieve that goal. First, to obtain the activity prediction for the whole video, we simply compute the mean over all video clips in the video. We then take the class with maximum predicted probability as the predicted class.

@@

\subsection{Dataset}
For all our experiments we use the dataset provided in the ActivityNet Challenge 2016~\cite{caba2015activitynet}. This dataset contains 640 hours of video and 64 million frames. The ActivityNet dataset is composed of untrimmed videos, providing temporal annotations for the given ground truth class labels. The dataset is split in 50\% for training, 25\% for validation and 25\% for testing.

\subsection{Training}
We train the network described in Section~\ref{sec:architecture} with the negative log likelihood loss, assigning a lower weight to background samples to deal with dataset imbalance (see Equation~\ref{eq:loss}).

@@

where $q$ is the predicted probability distribution and $p$ the ground truth probability distribution. In our experiments, we set $\rho$ = .3.

The network was trained with a 256 batch size and 20 \textit{timesteps} as input sequence.
The optimezer function used was the RMSprop~\cite{dauphin2015rmsprop} with a learning rate set to $10^{-5}$. In total the network was trained with 100 epochs.

% AMAIA: Aquí tambe hauries d'afegir el numero de iteracions (o epochs), learning rate, optimizer, batch size. Si el tens, el temps d'entrenament en la GPU que entrenessis.
% ALBERTO: FET

@@

The evaluation metric follow the metric used in ActivityNet Challenge to regard classification and temporal activity localization as retrieval probelm and evaluate the mAP.
For temporal localization, a prediction is masrked correct only when it has the correct category and has IoU with ground truth instance larger than $0.5$.

\begin{table}[h]
    \parbox{.45\linewidth}{
        \centering
        \begin{tabular}{l|cc}
        \textbf{Architecture} & \textbf{mAP} & \textbf{Hit@3} \\
        \hline
        3 x 1024-LSTM & 0.5635 & 0.7437 \\
        2 x 512-LSTM & 0.5492 & 0.7364 \\
        1 x 512-LSTM & \bf0.5938 & \bf0.7576 \\
        \end{tabular}
        \vspace{.5cm}
        \caption{Results for classification task comparing different architectures.}
        \label{table:classification_by_architecture}
    }
    \hfill
    \parbox{.45\linewidth}{
        \centering
        \begin{tabular}{l|ccc}
        \textbf{$\gamma$} & \textbf{$k=0$} & \textbf{$k=5$} & \textbf{$k=10$} \\
        \hline
        0.2 & 0.20732 & \bf0.22513 & 0.22136 \\
        0.3 & 0.19854 & 0.22077 & 0.22100 \\
        0.5 & 0.19035 & 0.21937 & 0.21302 \\
        \end{tabular}
        \vspace{.2cm}
        \caption{mAP with an IOU threshold of $0.5$ comparing between values of $k$ and $\gamma$ on post-processing.}
        \label{table:detection_postprocessing_comparison}
    }
\end{table}

The first results are related with the network architecture. It has been tested with different number of layers and different number of cells having all the layers the same number of cells. The results can be found in Table~\ref{table:classification_by_architecture}.
This results indicate that all the networks presented high learning capacity over the data ans it was obeseved a little over-fitting with the deeper architectures, obtainng the best results with a single layer of 512-LSTM cells.

Once the architecture has been chosen to obtain the best performance in the classification task, some experiments with different values of $k$ and $\gamma$ for the post-processing were performed.
At Table~\ref{table:detection_postprocessing_comparison} there is the results for the temporal activity localization task, where can be seen that the effect of a mean filter to smooth helps improving the localization.

% \begin{table}[h]
% \begin{center}
% \begin{tabular}{|l|c|c|}
% \hline
% \textbf{Architecture} & \textbf{mAP} & \textbf{Hit@3} \\
% \hline\hline
% 3 x 1024-LSTM & 0.5635 & 0.7437 \\
% 2 x 512-LSTM & 0.5492 & 0.7364 \\
% 1 x 512-LSTM & \bf0.5938 & \bf0.7576 \\
% \hline
% \end{tabular}
% \end{center}
% \caption{Results for classification task comparing different deep architectures. All values with
%          only video features on the validation dataset.}
% \label{table:classification_by_architecture}
% \end{table}
%
% \begin{table}[h]
% \begin{center}
% \begin{tabular}{|l|c|c|c|}
% \hline
% \textbf{$\gamma$} & \textbf{$k=0$} & \textbf{$k=5$} & \textbf{$k=10$} \\
% \hline
% \textbf{0.2} & 0.20732 & \bf0.22513 & 0.22136 \\
% \textbf{0.3} & 0.19854 & 0.22077 & 0.22100 \\
% \textbf{0.5} & 0.19035 & 0.21937 & 0.21302 \\
% \hline
% \end{tabular}
% \end{center}
% \caption{mAP with an IOU threshold of $0.5$ over validation dataset. Here there is a comparison
% between values of $k$ and $\gamma$ on post processing.}
% \label{table:detection_postprocessing_comparison}
% \end{table}
% \end{multicols}

% \begin{figure}
% \centering
% \subfloat[][]{
% \begin{table}[h]
% \begin{center}
% \begin{tabular}{|l|c|c|}
% \hline
% \textbf{Architecture} & \textbf{mAP} & \textbf{Hit@3} \\
% \hline\hline
% 3 x 1024-LSTM & 0.5635 & 0.7437 \\
% 2 x 512-LSTM & 0.5492 & 0.7364 \\
% 1 x 512-LSTM & \bf0.5938 & \bf0.7576 \\
% \hline
% \end{tabular}
% \end{center}
% \caption{Results for classification task comparing different deep architectures. All values with
%          only video features on the validation dataset.}
% \label{table:classification_by_architecture}
% \end{table}
% }
% \qquad
% \begin{table}[h]
% \begin{center}
% \begin{tabular}{|l|c|c|c|}
% \hline
% \textbf{$\gamma$} & \textbf{$k=0$} & \textbf{$k=5$} & \textbf{$k=10$} \\
% \hline
% \textbf{0.2} & 0.20732 & \bf0.22513 & 0.22136 \\
% \textbf{0.3} & 0.19854 & 0.22077 & 0.22100 \\
% \textbf{0.5} & 0.19035 & 0.21937 & 0.21302 \\
% \hline
% \end{tabular}
% \end{center}
% \caption{mAP with an IOU threshold of $0.5$ over validation dataset. Here there is a comparison
% between values of $k$ and $\gamma$ on post processing.}
% \label{table:detection_postprocessing_comparison}
% \end{table}
%
% \end{figure}

In the Figure~\ref{fig:example_results} is shown some examples of the temporal localization prediction for some instances of the dataset. It can be seen how the predictions are continuous and smooth and mostly overlaps with the ground truth instances.

\begin{figure}[ht]
\centering
\includegraphics[width=1\linewidth]{img/activity_temporal_localization_8.png}
\includegraphics[width=1\linewidth]{img/activity_temporal_localization_11.png}
\includegraphics[width=1\linewidth]{img/activity_temporal_localization_20.png}
\includegraphics[width=1\linewidth]{img/activity_temporal_localization_35.png}

\caption{Global architecture of the proposed pipeline.}
\label{fig:example_results}
\end{figure}
% AMAIA: Crec que amb una secció on expliquis tots els experiments ja n'hi ha prou. Li he posat el nom de 'Results'

% AMAIA: Si aqui vols parlar del bias towards sport categories, hauries d'afegir la figura que tens que ho demostra.
% ALBERTO: crec que potser una segona figura no hi cab...posar-la a annexes potser???
\section{Conclusion}

We propose a simple pipeline which can be able to perform in classification and temporal localization on activities on videos. Moreover, the flexibility of this sequence to sequence prediction allows the proposed network to face tasks where more than a single activity is present in the video.

In the future, we would like to extend the work training all the pipeline end-to-end, including the 3D-CNN and improve the accuracy accuracy obtained.
